\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-03-06 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
Update Manuscript_ViscousInstabilities_KC.tex
Section 1 and VFinger

\medskip
Due to complex and/or heterogeneous geological formations, multiphase flows through porous media play an important role in understanding the afformentioned operations. Modeling of multiphase flow, is an enormous technical and computational challenge. Starting from important contributions on the field of multiphase flows from \citet{wooding_1976}, \citet{homsy_1987} and \citet{adler_1988}, in the research area of fluid mechanics, there are the instabilities, also know as flow instabilities as these, first described by \citet{saffman_1958}. The first distinction should be made between microscopic and macroscopic instabilities. Microscopic instabilities are used for phenomena occurring locally, for example, Kelvin-Helmholtz and Saffman-Taylor instabilities \citep{saffman_1959}. On the other hand the macroscopic instabilities involves two-phase flow system \citep{bottoni_1992}. In this paper, the test-cases taht will follow focus on macroscopic instabilities and phenomena. A local flow instability that is the focus of this work is due to viscous and stress forces and often named as viscous instabilities or \textit{viscous fingering}.

@@

\medskip
In viscous fingering, a less viscous fluid pushes a more viscous fluid, in a thin rectangular cell (known as Hele-Shaw cell). \citet{howison_2000}  Under the assumption that fluids remain immiscible along an interface, surface tension plays an important role in determining the shape and progress of the fingers. For homogeneous domains fingers will start to develop when surface tension acting on the interface, is opposing towards the change of shape of the interface thus the interface becomes unstable and will collapse, taking a curved shape. In heterogeneous environments this process may also be triggered by permeability differences between smaller regions. The higher the velocity of the injected fluid, the less wide (tip-splitting behaviour) the finger will be. In addition, pressure differences %(eq.\ref{eq:pressure_dif}) 
acting on the interface will produce a net pressure force,

@@

\noindent this is also know as the Young-Laplace equation, a relation describing the capillary pressure difference across the interface between two fluids. With $\Delta P$ denoting the pressure difference, $\hat{n}$ the unit normal vector out of the surface, and $\gamma$ the surface tension. As mentioned in Section \ref{section:intro}, an unfavourable mobility-ratio in a water-oil system will have the following definition, see eqn.\ref{eq:MR}.

\begin{equation}
 MR \; = \frac{k_{rw} \; \mu_{w}}{k_{ro} \; \mu_{o}} 
\label{eq:MR}
\end{equation}

\noindent Fingering will occur in immiscible displacements when the viscocity ratio is greater than unity. %As surface tension becomes weak, the front of the steady finger is susceptible to a viscous-fingering instability by the basic mechanism associated with a less viscous fluid displacing a more viscous one. After a split, each of the new lobes of the finger is stable as a result of their being thinner than the finger from which they split. 
As surface tension becomes weak, the front is stressed and goes under a viscous-fingering instability. The front will collapse and will start splitting into new lobes of fingers. One of these fingers will eventually outgrow the others and will then spread to occupy the appropriate width of the cell. In the process, it reaches a width that is again unstable to splitting, and the pattern repeats. Thus, surface tension plays an essential dual role; it must be weak enough for the tip front to be unstable, but it is also the physical force causing the \textit{spreading} and ensuing repeated branching.

>>>>

\medskip
Due to complex and/or heterogeneous geological formations, multiphase flows through porous media play an important role in understanding the afformentioned operations. Modeling of multiphase flow, is an enormous technical and computational challenge. Starting from important contributions on the field of multiphase flows from \citet{wooding_1976}, \citet{homsy_1987} and \citet{adler_1988}, in the research area of fluid mechanics, there are the instabilities, also know as flow instabilities as these, first described by \citet{saffman_1958}. The first distinction should be made between microscopic and macroscopic instabilities. Microscopic instabilities are used for phenomena occurring locally, for example, Kelvin-Helmholtz and Saffman-Taylor instabilities \citep{saffman_1959}. On the other hand the macroscopic instabilities involves two-phase flow system \citep{bottoni_1992}. In this paper, the test-cases will focus on macroscopic flow instabilities due to viscous and stress forces also often named as viscous instabilities or \textit{viscous fingering}.

@@

\medskip
The physics as well as the main ide of the viscous fingering is the described experimentaly by the Hele-Shaw cell. Two layers really close to each other they they form an infinite lenght while a less viscous fluid pushes a more viscous one in between this rectangular cell. \citet{howison_2000}  Under the assumption that fluids remain immiscible along an interface, surface tension plays an important role in determining the shape and progress of the fingers. For homogeneous domains fingers will start to develop when surface tension acting on the interface, is opposing towards the change of shape of the interface thus the interface becomes unstable and will collapse, taking a curved shape. In heterogeneous environments this process may also be triggered by permeability differences between smaller regions. The higher the velocity of the injected fluid, the less wide (tip-splitting behaviour) the finger will be. In addition, pressure differences %(eq.\ref{eq:pressure_dif}) 
acting on the interface will produce a net pressure force, expressed by the Young-Laplace equation,

@@

\noindent a relation describing the capillary pressure difference across the interface between two fluids. In Eq. \ref{eq:pressure_dif} $\Delta P$ denotes the pressure difference, $\hat{n}$ is the unit normal vector out of the surface, and $\gamma$ the surface tension. As mentioned in Section \ref{section:intro}, the mobility-ratio in a water-oil system can be defined as

\begin{equation}
 MR \; = \frac{K_{rw} \; \mu_{w}}{K_{ro} \; \mu_{o}} 
\label{eq:MR}
\end{equation}

\noindent Fingering occurs in immiscible displacements when the viscocity ratio is greater than unity. %As surface tension becomes weak, the front of the steady finger is susceptible to a viscous-fingering instability by the basic mechanism associated with a less viscous fluid displacing a more viscous one. After a split, each of the new lobes of the finger is stable as a result of their being thinner than the finger from which they split. 
As surface tension becomes weak, the front is stressed and undergoes a viscous-fingering instability. The front collapses and starts splitting into new lobes of fingers. One of these fingers eventually outgrows the others and then spreads to occupy the appropriate width. In the process, it reaches a width that is again unstable resulting in splitting, and the cycle repeats. Thus, surface tension plays an essential dual role; it must be weak enough for the tip front to be unstable, but it is also the physical force causing the \textit{spreading} and ensuring repeated branching.
